Are populations like a circuit? The relationship between isolation by distance and isolation by resistance

### 2018-08-31 isolation_by_coalescence.tex
moar words

\paragraph{Process noise}
The mean coalescence times of lineages that we compute analytically
are exact for large, randomly mating populations,
but the word ``mean'' indicates an average across several levels of randomness:
and observed, empirical mean estimated using a sample of genotyeps
will differ slightly depending on both the individuals sampled (``sampling noise'')
and the stochasticity of the population history itself (``process noise'').
For more discussion of this latter contribution in other contexts, see
\citet{waples}, \citet{wakeley}, or \citet{ralph_empirical}.
In populations living across continuous geography,
both sources of noise can be substantially larger due to randomness in spatial locations.
A hint of this is visible in Figure \ref{fig:ind_locs_5x3b_1},
where we see that the distribution of individuals across the landscape
is fairly patchy.
This might make inference using a discretization of continuous space considerably more difficult.

@@

Although commute and coalescence times can be quite different,
we show that they are equal (with a particular choice of local diversity)
if populations are arranged on a ring or torus
with symmetric migration.
The similarity of these models to rectangular grids
may explain why in other work, commute time methods have shown relatively good fit to data
simulated with symmetric migration rates.

@@

\paragraph{Asymmetric dispersal}
(Review of asymmetric situations.)

@@

\paragraph{Landscapes or discrete demes?}
Other inference methods \citep{wilson2003bayesian,beerli1999maximum,beerli2010unified,hey2007integration}
work with coalescence inference, but only with a small number of loci and populations,
because of computational efficiency.
We're somewhere in between.
\plr{XXX}

>>>>

\paragraph{Process noise}
The mean coalescence times of lineages that we compute analytically
are exact for large, randomly mating populations,
but the word ``mean'' indicates an average across several levels of randomness:
and observed, empirical mean estimated using a sample of genotyeps
will differ slightly depending on both the individuals sampled (``sampling noise'')
and the stochasticity of the population history itself (``process noise'').
For more discussion of this latter contribution in other contexts, see
\citet{wakeley2012genealogies}, \citet{waples2009modeling}, or \citet{ralph2015empirical}.
In populations living across continuous geography,
both sources of noise can be substantially larger due to randomness in spatial locations.
A hint of this is visible in Figure \ref{fig:ind_locs_5x3b_1},
where we see that the distribution of individuals across the landscape
is fairly patchy.
This might make inference using a discretization of continuous space considerably more difficult.

@@

Although commute and coalescence times can be quite different,
as also noted by \citet{mcrae2006isolation}, 
they are equal (with a particular choice of local diversity)
if populations are arranged on a ring or torus
with symmetric migration.
The similarity of these models to rectangular grids
may explain why in other work, commute time methods have shown relatively good fit to data
simulated with symmetric migration rates.

@@

\paragraph{Other methods}
Other methods that use resistance distance have substantial advantages over the method we present here.
Circuitscape \citep{mcrae2008using} produces maps of much higher resolution 
than is presently possible with coalescent times.
\citet{hanks2013circuit} connects the problem to intrinsic conditional autoregressive models, 
providing a Bayesian model of an underlying Gaussian Markov random field.
EEMS \citep{petkova2014visualizing} integrates over possible tesselations of the region
to obtain a landscape at the best possible resolution supported by the data.

Although we present a new method, based on coalescence times,
our focus is on understanding the underlying modeling issues
rather than replacing these other tools.
Most promisingly, a very recent method, MAPS \citep{alasadi2018estimating}
is based on EEMS, but instead of resistance distances,
uses pairwise sharing of long genomic segments (``IBD segments'', \citet{browning2013improving}),
which have been shown to carry substantial information about recent demography
\citep{ralph2013geography,palamara2013inference,browning2015accurate}.
Furthermore, the likelihood model underlying MAPS is based on coalescent theory,
and thus should not be subject to the drawbacks of resistance distance.
However, application requires fairly dense genotyping and a good genetic map,
something not available for many species.

Other widely-used inference methods, such as
MIGRATE \citep{beerli1999maximum,beerli2010unified},
BayesAss \citep{wilson2003bayesian}, 
and IMa \citet{hey2007integration}
use likelihoods derived from a coalescent model --
and use much more of the data than we do.
However, because they use locus-specific data rather than an average across loci,
these methods can only be applied to a relatively small number of populations and loci,
because of computational efficiency.

\paragraph{Asymmetric dispersal}
We have found that asymmetric rates of lineage movement
are most likely to produce qualitative differences between coalescence times and resistance distances,
and thereby mislead models based on resistance distance.
This asymmetry occurs if one region receives a greater proportion of each new generation as migrants
from another population than that other population does from it.
This is expected, for instance,
if organismal dispersal is biased by wind or water currents \citep{gaines2003avoiding,morrissey2009maintenance},
or in the presence of source-sink dynamics \citep{dias1996sources,lenormand2002limits}.
XXX MORE

@@
